---
title: "Homework 6"
author: "Raul Torres Aragon"
date: "11/01/2022"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
options(scipen=999)
load("hw6_objects.RData")
```  

## 1.  

Table 1 reproduce data from Altham (1991) of counts of T4 cells/mm3 in blood samples from 20 patients 
in remission from Hodgkin’s disease and 20 other patients in remission from disseminated malignancies. 
The question of interest here is: Is there a difference in the distribution of cell counts between the two diseases? 
A quantitative assessment of any difference is also required.  


## (a)  
Carry out an exploratory data analysis and provide a summary of the two distributions.  

The following boxplots show the distribution of count of cells by disease. 
Min, max, and means are displayed on the figure. Notice that individuals with Hodgkin's show 
a higher mean count of cells on min, mean, max. On the surface it may seem like this difference 
is statistically significant, but it could very well be driven by the outlier at 2,415.  


```{r, echo=FALSE, fig.align='center', out.width="70%"}
boxplot(data.frame("Non_Hodgkins" = nh, "Hodgkins" = ho),
        main = "Cell counts\n between two diseases",
        horizontal = TRUE)
text(c(mean(nh), min(nh), max(nh), min(ho), max(ho), mean(ho)),  
     c(rep(0.8,3), rep(1.8,3)),
     c(mean(nh), min(nh), max(nh), min(ho), max(ho), mean(ho)), 
     pos = 1, cex = 0.6
     )  
points(means, c(1,2), col="steelblue",pch=1, cex=2)
```  

## (b)  
We now obtain 90% confidence intervals for differences of means between distributions on 
the original scale, log transformed, and square root transformed.  

Recall the equation for a 90% confidence interval for a difference in means is:  

$$
\bigg((\mu_{1}-\mu_2) -\Phi(.05) \sqrt{\frac{\sigma^2_{1}}{n_1}+\frac{\sigma^2_{2}}{n_2}},
(\mu_{1}-\mu_2) -\Phi(.05) \sqrt{\frac{\sigma^2_{1}}{n_1}+\frac{\sigma^2_{2}}{n_2}}\bigg)
$$  
we obtain the variance (after applying the transformation) of each group count via bootstrap, 
making sure the sampling is separate for the Hodgkin's and the Non Hodgkin's groups.  

The resulting confifence intervals are:  

* original scale: (`r round(CI_orig, 2)`)  
* log scale: (`r round(CI_log, 2)`)  
* sqrt scale: (`r round(CI_sqrt, 2)`)  

The considerations have to do with our belief of unequal variance in the data. For example, 
the log transform can ameliorate heteroscedasticity. If we don't believe heterscedasticity is 
present in our sample, then we can just assume homoscedasticity and proceed with the original 
scale.  

## (c)  
We want to know whether the means across the two groups are different. To test that with 
different models, we first pull the data to create one dependent variable with all counts. 
Then, we create a dummy variable to distinguish counts from Hodgkin's patients and non Hodgkin's 
patients. Finally, we fit three models: (1) poisson, (2) Gamma, and (3) inverse Gaussian, using 
their respective canonical forms: $X\beta = log(\mu)$, $X\beta = -\mu^{-1}$, and 
$X\beta = \frac{1}{\mu^2}$.  

Here are the results in their respective link scale, and then in the original scale:  

```{r, echo=FALSE}
knitr::kable(res_tab, 
             caption = "Results of three models for count of cells by whether patient has Hodgekins. (link scale)",
             digits = 4) 
```  

We now present the same results but "undoing" the link function to recover the original scale.  
In other words,  
For Poisson, we exponentiate the coefficients because $\exp(X\beta) = \mu$  
For Gamma, we take the negative inverse because $-\frac{1}{X\beta} = \mu$  
For inverse Gaussian we take the positive squared inverse because $\frac{1}{\sqrt{X\beta}} = \mu$.  



```{r, echo=FALSE}
knitr::kable(res_tab_undone, 
             caption = "Results of three models for count of cells by whether patient has Hodgekins. (original scale)",
             digits = 4) 
```  
  
Recall that the intercept houses the mean count for those without Hodgkin. $\beta_1$ is the 
difference in cell count between those with Hodgkin's and those without it. Thus, testing whether $\beta_1 = 0$ 
is synonymous with testing whether there is a statistically significant difference in mean counts. Hence:  
$H_0: \beta_1 = 0$ while $H_1: \beta_1 \neq 0$.    

## (d)  

Using the asymptotic distribution of the MLE, that is  

$$
\mathcal{I}(\hat{\beta})^{1/2}(\hat{\beta} - \beta) \to_d \text{N}_2(0,I_2)
$$  
we give 90% confidence intervals for each parameter under each of the distributional assumptions. (Poison imposes a log distribution, Gamma imposes a negative inverse, and Inverse Gaussian imposes a squared inverse.  

Recall, the information matrix is obtain from `glm` by `solve(vcov(model))`. 
It's then easy to construct asymptotic confidence intervals with as follows:  

$$
\beta_i \pm \Phi(0.95)\times \sqrt{I_{i,i}^{-1}}
$$  

```{r, echo = FALSE}
knitr::kable(CI_tab, caption = "90% Confidence interval for each parameter under three models")
```

Based on the above results, I would conclude that, under Poisson model, there is a 
significant difference between the mean count of cells among patients with and without Hodgkin's.  
However, under the Gamma and the Inverse Gaussian model, I would not be so quick to do so given that 
zero is encompassed by the confidence intervals.  


## 2.  
The data in Table 2, taken from Wakefield et al. (1994), were collected following the 
administration of a single 30mg dose of the drug Cadralazine to a cardiac failure patient. 
The response $y_i$ represents the drug concentration at time $x_i$. 
The most straightforward model for these data is 
$$
log(y_i) = \mu(\beta) + \epsilon_i = log\bigg[\frac{30}{V}\exp(-k_ex_i)\bigg] + \epsilon_i
$$  
```{r, echo=FALSE}
knitr::kable(data.frame(x,y))
```

## (a)  

We now obtain expressions for  
i) Log-likelihood function:  

Note that we assume $log(y_i) \sim Normal \implies y_i \sim logNormal$, thus we start with the 
lognormal distribution and compute the loglikelihood function from there. 

$$
l(\boldsymbol{\beta}, \sigma^2) = 
-\frac{n}{2} log(2\pi\sigma^2) - \sum_i^n log(y_i)-\frac{1}{2\sigma^2}\sum_{i=1}^n[log(y_i)-log(30/V)+K_ex_i]^2
$$  


ii) Score function  
The Score function is a vector-valued function of length = 3 given $\beta_0, \beta_1$, and $\sigma^2$.  
$$
\mathcal{S}(\boldsymbol{\beta},\sigma^2) =  
\bigg[\frac{\partial l}{\partial \beta_0},  
      \frac{\partial l}{\partial \beta_1}, 
      \frac{\partial l}{\partial \sigma_2},\bigg]
$$  
$$
= \bigg[
\frac{\sum_{i=1}^n(log(y_i)-\beta_0-\beta_1x_i)}{\sigma^2} ,
\frac{\sum_{i=1}^n(log(y_i)-\beta_0-\beta_1x_i)x_i}{\sigma^2}, 
\frac{-n}{\sigma^2} + \frac{\sum_{i=1}^n(ln(y_i)-\beta_0-\beta_1x_i)^2}{(\sigma^2)^2}
\bigg]  
$$


where $\beta_0 = log(30/V)$, $\beta_1 = -K_e$, and $\sigma^2 = \sigma^2$.  

iii) Expected information matrix $\mathcal{I}(\boldsymbol{\beta},\sigma^2)$  
$$
\mathcal{I}(\boldsymbol{\beta},\sigma^2) = -E\bigg( 
\begin{bmatrix}
\frac{\partial l}{\partial \beta_0 \beta_0} & 
\frac{\partial l}{\partial \beta_0 \beta_1} & 
\frac{\partial l}{\partial \beta_0 \sigma^2} \\
\frac{\partial l}{\partial \beta_1 \beta_0} & 
\frac{\partial l}{\partial \beta_1 \beta_1} & 
\frac{\partial l}{\partial \beta_1 \sigma^2} \\
\frac{\partial l}{\partial \sigma^2 \beta_0} & 
\frac{\partial l}{\partial \sigma^2 \beta_1} & 
\frac{\partial l}{\partial \sigma^2 \sigma^2}
\end{bmatrix} \bigg)= 
$$  
$$
\begin{bmatrix}
\frac{n}{\sigma^2}E[log(y_i)] & 
\frac{\sum_{i=1}^n\sum_{i=1}^nx_i}{\sigma^2}E[log(y_i)] & 
\frac{n}{(\sigma^2)^2}E[log(y_i)] \\
\frac{\sum_{i=1}^n\sum_{i=1}^nx_i}{\sigma^2}E[log(y_i)] & 
\frac{nE[log(y_i)]}{\sigma^2} \sum_{i=1}^n x_i& 
\frac{\sum_{i=1}^n\sum_{i=1}^nx_i}{(\sigma^2)^2}E[log(y_i)] \\
\frac{n}{(\sigma^2)^2}E[log(y_i)] & 
\frac{\sum_{i=1}^n\sum_{i=1}^nx_i}{(\sigma^2)^2}E[log(y_i)]  & 
\frac{2}{(\sigma^2)^3} [nE(ln(y_i)^2) + n\beta_0E(log(y_i)) +\beta_1E(log(y_i))\sum_{i=1}^nx_1]
\end{bmatrix}
$$  
where $E[log(y_i)]=log(\exp\{\mu_y - \sigma^2/2\})=\mu_y - \sigma^2/2$.


## (b)  
Using `optim` to numerically optimize the log-function for $V, -K_e, \sigma^2$, we get MLEs and their asymptotic 95% confidence intervals:  

```{r, echo=FALSE}
knitr::kable(Q2b_tab, caption = "MLEs for V, Ke, and sigma^2")
```

## (c)  
We now plot the data, along with the fitted curve.  

```{r, echo=FALSE, fig.align='center', out.width="70%"}
plot(y~x, main = "Cadralazine concentration at 8 given times", xaxt='n')
lines(exp(b0-b1*x_grid), col = "steelblue", lty=1, lwd=3)
axis(1,at=x,labels=x)
```

## (d)  
Using residuals, we examine the appropriateness of the assumptions of the above model. 
And answer the question: Does the model seem reasonable for these data?  

To examine the assumption of equal variance (homoscedasticty) we plot residuals against 
fitted values.  
```{r, echo=FALSE, fig.align='center', out.width="70%"}
par(mfrow = c(1,2))
plot(r~yhat, 
     main = "residual-fitted plot", 
     xlab = "time", ylab = "residual",
     xaxt="n")
axis(1,at=x,labels=x)
plot(std_r~yhat,
     main = "std residuals-fitted plot",
     ylab = "std residuals")
```

Next, we examine the assumption of no correlation among error terms. This time, we plot 
residuals against lag(1) residuals. 
```{r, echo=FALSE, fig.align='center', out.width="70%"}
acf(r, main = "Autocorrelation plot for residuals")
```  

Finally, given our small sample, we want to determine whether the residuals are normally 
distributed. The following QQ-plot compares the distribution of our residuals with theoretical 
quantiles of a normal distribution.    

```{r, echo=FALSE, out.width="70%"}
qqnorm(r, pch = 1, frame = FALSE)
qqline(r, col = "steelblue", lwd = 2)
```

The model does not seem to be a great fit for these data because it t consistently 
underestimates the cell count as shown in (c). Furthermore, the residual plots seem to 
suggest variance is not constant. There does seem to be a megaphone pattern growing at 
later times (ignoring the outlier). There does not seem to be correlation between residual 
terms, though, the sample size may be too small to get a fairer assessment. The correlation 
between $r_t$ and $r_{t-1}$ is `r round(cor(r[2:8], dplyr::lag(r)[2:8]),2)`, which does not seem very 
strong. The residuals do seem to be normally distributed, so with the exception of 
equal variance, the LR assumptions seem to hold.  

## (e)  
The clearance $Cl = V \times K_e$ and elimination half-life $x_{1/2} = log(2)/K_e$ are 
parameters of interest in this experiment.  
We find the MLEs of these parameters along with asymptotic 95% confidence intervals.  

$\hat{Cl} = \hat{V}\hat{K_e}$ and 
$\hat{x}_{1/2} = log(2)/\hat{K_e}$  

To construct a confidence interval, we need the asymptotic variance, which we obtain via 
Delta Method.  
Notice Cl is a function of two parameters, namely V and Ke. By the Delta Method,  

$$
var(Cl) = \nabla l(V, K_e)^T I(V,K_e)^{-1}\nabla l(V,K_e)
$$  
where $\nabla l(K_e, V)=$  

$$
\bigg[
-\frac{K_e}{\sigma^2} \sum_{i=1}^n(log(y_i) - log(30/V) + K_ex_i), 
-\frac{1}{\sigma^2V} \sum_{i=1}^n(log(y_i) -log(30/V) + K_ex_i)
\bigg]
$$  
and $\mathcal{I}(K_e, V)=$  
$$
-E\bigg(\begin{bmatrix}
\frac{\partial l}{\partial K_e K_e} & -\frac{\partial l}{\partial K_eV} \\
-\frac{\partial l}{\partial VK_e} & \frac{\partial l}{\partial VV}]
\end{bmatrix} \bigg)
$$  

$$
-E\bigg(
\begin{bmatrix}
\frac{1}{\sigma^2}\sum_{i=1}^n[-log(y_i)+log(30/V) -2K_ex_i] & -\frac{-K_e}{\sigma^2V} \\
-\frac{-K_e}{\sigma^2V} & \frac{1}{\sigma^2V^2}\sum_{i=1}^n[K_ex_i-log(30/V)+log(y_i)-1]
\end{bmatrix} \bigg)
$$  
$$
\begin{bmatrix}
-\frac{n\mu}{\sigma^2} - \frac{n}{2} & 0 \\
0 & -\frac{1}{V^2}(\frac{n\mu}{\sigma^2} + \frac{n}{2})
\end{bmatrix}
$$

and for the half-life we have (by the single variable Delta Method):    
$$
var(x_{1/2}) = \bigg(\frac{\partial l(K_e,V)}{\partial K_e}\bigg)^2 \times \mathcal{I}(K_e,V)_{[1,1]}^{-1}
$$  
$$  
\bigg(-\frac{K_e}{\sigma^2} \sum_{i=1}^n(log(y_i) - log(30/V) + K_ex_i)\bigg)^2 \times \frac{1}{-\frac{n\mu}{\sigma^2} - \frac{n}{2}}
$$  

where $\mu = \exp\{log(30/V) \exp(-K_ex_i))\}$ since $y_i \sim \text{lognormal}(log(30/V)\exp\{-K_ex_i\}, (\exp(\sigma^2)-1)\exp(2[log(30/V)\exp\{-K_ex_i\}]\sigma^2))$.  

With the MLE estimates and the variance of each estimate, we can construct confidence 
interval as follows:  
$$
\hat{\theta}_{MLE} \pm \Phi(0.95)\times \sqrt{var(\hat{\theta})}
$$  
where $\hat{\theta}$ is our estimate.  

```{r, echo=FALSE}
mytab <- data.frame(param = round(c(Cl_mle, xhalf_mle),1),
                    conf_int = c(CI_cl, CI_xhalf))
colnames(mytab) <- c("MLE estimate", "95% conf interval")
row.names(mytab) <- c("clearance","half-life")
knitr::kable(mytab, caption = "MLE for two parameters of interest")

```

* Cl = `r CI_cl`  
* $x_{1/2}$ = `r CI_xhalf`  

## Apendix  
```{}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# student: Raul
# course: STAT 570
# assignment: hw6
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

rm(list=ls())
library(tidyverse)
library(dplyr)


nh <-c(375,	375,	752,	208,	151,	116,	736,	192,	315,	1252,	675,	700,	440,	771,	
       688,	426,	410,	979,	377,	503)
ho <-c(396,	568,	1212,	171,	554,	1104,	257,	435,	295,	397,	288,	1004,	431,	795,	
       1621,	1378,	902,	958,	1283,	2415)


#~~~~~~~~~~~#
# Problem 1 # 
#~~~~~~~~~~~#

## (a)
means <- c(mean(nh), mean(ho))
boxplot(data.frame("Non_Hodgkins" = nh, "Hodgkins" = ho),
        main = "Cell counts\n between two diseases",
        horizontal = TRUE)
text(c(mean(nh), min(nh), max(nh), min(ho), max(ho), mean(ho)),  
     c(rep(0.8,3), rep(1.8,3)),
     c(mean(nh), min(nh), max(nh), min(ho), max(ho), mean(ho)), 
     pos = 1, cex = 0.5)  
points(means, c(1,2), col="red",pch=1, cex=2)


## (b)  
get_bootstrap_var <- function(x, N) {
    B <- rep(0, N) 
    for(i in 1:N) {
        B[i] <- sample(x, size = length(x), replace = TRUE) |> var()
    }
    var <- mean(B)
    var
}
N <- 1e3
get_CI <- function(x1, x2, level = .90, scale = "orig") {
  if(scale == "log") {
    x1 <- log(x1); x2 <- log(x2)
  } else if(scale == "sqrt") {
    x1 <- sqrt(x1); x2 <- sqrt(x2)
  } else {
    TRUE
  }
  var1 <- get_bootstrap_var(x1, N=1e3)
  var2 <- get_bootstrap_var(x2, N=1e3)
  Phi <- abs(qnorm((1-level)/2, mean = 0, sd = 1))
  ci_lo <- mean(x1)-mean(x2) - Phi*sqrt(var1/length(x1) + var2/length(x2))
  ci_hi <- mean(x1)-mean(x2) + Phi*sqrt(var1/length(x1) + var2/length(x2))
  return(c(ci_lo, ci_hi))
}

CI_orig <- get_CI(ho, nh, level = 0.90, scale = "orig")
CI_log  <- get_CI(ho, nh, level = 0.90, scale = "log")
CI_sqrt <- get_CI(ho, nh, level = 0.90, scale = "sqrt")


## (c)  
data <- data.frame("count" = c(nh, ho), 
                   "hodgkins" = c(rep(0,length(nh)), rep(1,length(ho))))

fit_pois <- glm(count~hodgkins, family=poisson(link="log"), data=data)
fit_gamm <- glm(count~hodgkins, family=Gamma(link="inverse"), data=data) 
fit_invg <- glm(count~hodgkins, family=inverse.gaussian(link="1/mu^2"), data=data)

coef_pois <- summary(fit_pois)$coefficients |> as.data.frame()
coef_gamm <- summary(fit_gamm)$coefficients |> as.data.frame()
coef_invg <- summary(fit_invg)$coefficients |> as.data.frame()

res_tab <- data.frame(
  "Intercept"    = c(coef_pois$Estimate[1], coef_gamm$Estimate[1], coef_invg$Estimate[1]),
  "SE_Intercept" = c(coef_pois$`Std. Error`[1], coef_gamm$`Std. Error`[1], coef_invg$`Std. Error`[1]),
  "Beta1"        = c(coef_pois$Estimate[2], coef_gamm$Estimate[2], coef_invg$Estimate[2]), 
  "SE_Beta1"     = c(coef_pois$`Std. Error`[2], coef_gamm$`Std. Error`[2], coef_invg$`Std. Error`[2])
  )
row.names(res_tab) <- c("Poisson","Gamma","InverseGaussian")

undo_gamma <- function(x) { -1/x }
undo_invg  <- function(x) { 1/sqrt(x) }

coef_pois_undone <- fit_pois$coefficients[,1] |> exp() |> as.data.frame()
coef_gamm_undone <- fit_gamm$coefficients[,1] |> undo_gamma() |> as.data.frame()
coef_invg_undone <- fit_invg$coefficients[,1] |> abs() |> undo_invg() |> as.data.frame()

res_tab_undone <- data.frame(
  "Intercept"    = c(coef_pois_undone[1,1], coef_gamm_undone[1,1], coef_invg_undone[1,1]),
  "Beta1"        = c(coef_pois_undone[2,1], coef_gamm_undone[2,1], coef_invg_undone[2,1]) 
)
row.names(res_tab_undone) <- c("Poisson","Gamma","InverseGaussian")



## (d)  
I_pois <- solve(vcov(fit_pois))
I_gamm <- solve(vcov(fit_gamm))
I_invg <- solve(vcov(fit_invg))

get_CI_model <- function(theta, I_jj) {
  r <- c(theta - qnorm(0.95)*sqrt(1/I_jj), theta + qnorm(0.95)*sqrt(1/I_jj))
  r
}

CI_b0_pois <- get_CI_model(coef_pois[1,1], I_pois[1,1])
CI_b1_pois <- get_CI_model(coef_pois[2,1], I_pois[2,2])

CI_b0_gamm <- get_CI_model(coef_gamm[1,1], I_gamm[1,1])
CI_b1_gamm <- get_CI_model(coef_gamm[2,1], I_gamm[2,2])

CI_b0_invg <- get_CI_model(coef_invg[1,1], I_pois[1,1])
CI_b1_invg <- get_CI_model(coef_invg[2,1], I_pois[2,2])


CI_tab <- data.frame("intercept" = c(paste0(round(CI_b0_pois, 3), collapse = " to "), 
                                     paste0(round(CI_b0_gamm, 3), collapse = " to "),
                                     paste0(round(CI_b0_invg, 3), collapse = " to ")),
                     "b1" = c(paste0(round(CI_b1_pois, 3), collapse = " to "), 
                              paste0(round(CI_b1_gamm, 3), collapse = " to "),
                              paste0(round(CI_b1_invg, 3), collapse = " to "))
           )

row.names(CI_tab) <- c("Poisson","Gamma","InverseGaussian")




#~~~~~~~~~~~#
# Problem 2 # 
#~~~~~~~~~~~#

x <- c(seq(from = 2, to = 10, by = 2), 24, 28, 32)

y <- c(1.63, 1.01, 0.73, 0.55, 0.41, 0.01, 0.06, 0.02)

## a
## b

my_likelihood <- function(theta) {
  b0 <- log(30/theta[1])
  b1 <- theta[2]
  n <- length(x)
  sigma2 <- theta[3]
  -n/2 * log(2*pi*sigma2) -sum(log(y)) -(1/(2*sigma2))*sum( (log(y) -b0+b1*x)^2 ) 
}
params <- optim(par = c(1,1,1), fn = my_likelihood, control = list(fnscale = -1))$par

V <-params[1]; Ke <- params[2]; sigma2 <- params[3]


# TO DO
I <- matrix(1, ncol=3, nrow=3)

q975 <- qnorm(.975, 0, 1)
ci_V <- c(V-q975*sqrt(I[1,1]), V+q975*sqrt(I[1,1])) |> round(2)
ci_Ke <- c(Ke-q975*sqrt(I[2,2]), Ke+q975*sqrt(I[2,2]))  |> round(2)
ci_sigma2 <- c(sigma2-q975*sqrt(I[3,3]), sigma2+q975*sqrt(I[3,3]))  |> round(2)

confidence_intervals <- c(paste0(ci_V, collapse = " to "),
                          paste0(ci_Ke,collapse = " to "),
                          paste0(ci_sigma2, collapse = " to "))

Q2b_tab <- data.frame(round(params,2), confidence_intervals)
row.names(Q2b_tab) <- c("V", "-Ke", "sigma2")
colnames(Q2b_tab) <- c("parameter", "95% conf interval")


## (c)  
#We now plot the data, along with the fitted curve.  
b0 <- log(30/V)
b1 <- Ke
x_grid <- seq(from = 2, to = 33, by = 1)
plot(y~x, main = "Cadralazine concentration at 8 given times", xaxt='n')
lines(exp(b0-b1*x_grid), lty=1, lwd=3)
axis(1,at=x,labels=x)

## (d)  
#Using residuals, we examine the appropriateness of the assumptions of the above model. 
#And answer the question: Does the model seem reasonable for these data?  
yhat <- exp(b0-b1*x)
r <- y - exp(b0-b1*x)
std_r <- (y - yhat) / sqrt(sigma2)
#hist(r, breaks = 13)


## (e)  
#The clearance Cl = V * K_e and elimination half-life x_{1/2} = log(2/K_e) are 
#parameters of interest in this experiment.  
#We find the MLEs of these parameters along with asymptotic 95% confidence intervals. 

Cl_mle <- V*Ke
xhalf_mle <- log(2)/Ke 
n <- length(x)

Mu <- exp(log(30/V*exp(Ke*x))) |> mean()

dKe <- sum(-log(y) + log(30/V) -2*Ke*x)/sigma2
dV <- sum(Ke*x -log(30/V) +log(y) -1)/(sigma2*V^2)
I_11 <- n*(Mu+sigma2/2)/sigma2
I_22 <- n*(Mu+sigma2/2)/(sigma2*V^2)
I_21 <- 0

grad_loglike <- matrix(c(dKe,dV), ncol=2, nrow=1)
FIM <- matrix(c(I_11, I_21, I_21, I_22), nrow=2, ncol=2)

var_Cl <- grad_loglike %*% solve(FIM) %*% t(grad_loglike)
var_xhalf <- dKe^2 * 1/I_22

CI_cl <- paste0(c(Cl_mle - qnorm(0.975)*sqrt(var_Cl), 
                  Cl_mle + qnorm(0.975)*sqrt(var_Cl)) |> round(3),
                collapse = " to ")

CI_xhalf <- paste0(c(xhalf_mle - qnorm(0.975)*sqrt(var_xhalf), 
                     xhalf_mle + qnorm(0.975)*sqrt(var_xhalf)) |> round(3),
                     collapse = " to ")
```







